---
title: "EBImage for cytopathology"
output:
  html_document:
    df_print: paged
---

# Conclusion
## steps
- blur the image 
- apply a threshold to turn cell into 'blobs'
- count the 'blobs'
## comments
- watershed function can not resolve two linked cells.


```{r include=FALSE}
#source("http://bioconductor.org/biocLite.R")
#biocLite("EBImage")
 library("EBImage")
```


```{r}
im <- readImage("papSmear_bg.jpg")
par(mfrow=c(1,2))
plot(im)
EBImage::display(im, method = "raster") # raster method means within R
```


 make the image brighter by mutiplying all the values by 2
 
```{r}
# convert image de gray scale
# im <- channel(im.c, "gray")
im2 <- im * 2
par(mfrow=c(1,2))
plot(im)
plot(im2)

```
 

```{r}
# gaussian blur
im.blur <- gblur(im, sigma= 0.01)
im2.blur <- gblur(im2, sigma = 1)
par(mfrow=c(1,2))
plot(im.blur)
plot(im2.blur)
```



# threshold using Otsu's method used to reduce of graylevel image to a binary image.
```{r}
# only gray images are accepted
# gives a threshold value using Otsu algorithm
otsu(im.blur)
#otsu(im2.blur)

#im.c2.g.blur.otsu <- im.c2.g.blur > otsu(im.c2.g.blur)

```

```{r echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
nuc = readImage(system.file("images", "nuclei.tif", package="EBImage"))
nuc_gblur = gblur(nuc, sigma = 5)
#display(nuc_gblur, method = "raster", all = TRUE)
threshold = otsu(nuc)
threshold_gblur <- otsu(nuc_gblur) 
threshold
threshold_gblur
#display(nuc_gblur > threshold, all =FALSE)
par(mfrow=c(1,2))
display(nuc > threshold, method = "raster", all = TRUE)
display(nuc_gblur > threshold, method = "raster", all = TRUE)

max(bwlabel(nuc > threshold))
max(bwlabel(nuc_gblur > threshold_gblur))

display(colorLabels(bwlabel(nuc > threshold)), method = "raster", all = TRUE)
display(colorLabels(bwlabel(nuc_gblur > threshold_gblur)), method = "raster", all = TRUE)

```





```{r}
im.blur.thr.cnt <- bwlabel(im.blur > otsu(im.blur))
N <- max(im.blur.thr.cnt)
N
```


```{r}
display(colorLabels(im.blur.thr.cnt), method= "raster") 
```


# R function to get label of cells
```{r}
getCelLab <- function(img, bright, sigma ){

im <- readImage(img)
im <- EBImage::channel(im, mode= "gray")
im2 <- im * bright
im.blur <- gblur(im, sigma)
plot(im.blur)
## filtering option 

im.blur.thr.cnt <- bwlabel(im.blur > otsu(im.blur))
#im.blur.the.cnt <- bwlabel(watershed( distmap(im.blur), 2 ))
#im.blur.thr.cnt <- bwlabel(adaptative_thre(31, "disc"))
#im.blur.thr.cnt <- propagate(seeds = im.blur, x = im.blur, lambda = 100)


N <- max(im.blur.thr.cnt)
display(colorLabels(im.blur.thr.cnt), method= "raster") 
return(N)
}
#par(mfrow=c(2,2))
getCelLab("papSmear.jpg", 2, 5)
getCelLab("papSmear_g.jpg", 2, 2)
getCelLab("papSmear_bg.jpg", 2, 0.1)
getCelLab("basale_inverse_gray.jpg", 2, 1)
```

```{r}
adaptative_thre <- function(size, shape){
disc = makeBrush(size, shape)
#disc <- fillHull(disc)
disc = disc / sum(disc)
offset = 0.05
im.blur_bg = filter2( im.blur, disc )
im.blur_th = im.blur > im.blur_bg + offset
return(im.blur_th)
}
plot(adaptative_thre(31, "disc"))
plot(adaptative_thre(9, "disc"))


```


# getCelLab with adaptative_thre
```{r}
getCelLab <- function(img, bright, sigma ){

im <- readImage("papSmear_bg.jpg")
im2 <- im * bright
im.blur <- gblur(im, sigma)
#im.blur.thr.cnt <- bwlabel(im.blur > otsu(im.blur))
im.blur.thr.cnt <- bwlabel(adaptative_thre(31, "disc"))
N <- max(im.blur.thr.cnt)
display(colorLabels(im.blur.thr.cnt), method= "raster") 
return(N)
}
#par(mfrow=c(2,2))
getCelLab("papSmear.jpg", 2, 5)
getCelLab("papSmear_g.jpg", 2, 2)
getCelLab("papSmear_bg.jpg", 2, 0.1)
```


# get graded magnetude image 
```{r include=FALSE}
library("imager")
```


```{r}
im <- load.image("papSmear_g.jpg")
dx <- imager::imgradient(im,"x")
dy <- imager::imgradient(im,"y")
grad.mag <- sqrt(dx^2+dy^2)
plot(grad.mag)
```

